### 2023-12-31 chapters/discussion.tex
Finished discussion on OODD for speech

\chapter[discussion]{Discussion}\label{chp:discussion}
% ~5 pages
%
% OUTLINE:
% - paper-hierarchical:
%   - Suitability of VAEs for representation learning (minimization of mutual information and sensitivity to implicit prior such as architecture)
% - paper-benchmarking
%   - Inferiority of probabilistic methods compared to self"=supervised learning. 
% -

% \todo[inline]{Unsupervised representation learning for medical applications. We tried BERT for stroke but no success. Medical coding state-of-the-art uses pretrained BERT, but only marginally improves upon simpler baselines.}

% \moretodo[inline]{Add to the discussion somewhere:
% I would try to embed a perspective for contemplation. Instead of us always talking about what the technology needs to improve on. Perhaps, it would also be prudent to ask an open-ended philosophical question: “While our requirements to the technology are in its place, one could ask, whether human beings have the ability to quantify uncertainty. In 2023, the prevalence of human errors in driving are much higher than self-driving vehicles.” <- Quick example. I'm sure you can find something better.}

@@

\moretodo{Adjust the below to include any discussion on out-of-distribution detection.}
Having examined VAEs and self"=supervised approaches for representation learning in \cref{part:unsupervised-uncertainty-estimation,part:unsupervised-speech-representation-learning}, in this discussion we will consider how VAEs might be enabled to learn representations more competitive with self"=supervised methods for downstream tasks. 
Furthermore, since the work presented in \cref{part:medical-applications} only had limited focus on uncertainty, we will also present and discuss the use of calibration techniques for the stroke recognition model, drawing connections to how such systems might be perceived and used in practice.

@@

\section{Out-of-distribution detection for speech}
%
% \moretodo[inline]{Out-of-distribution detection on speech?
%
% Outline:
% Intro paragraph drawing from OOD detection on images.
% Differences and unique challenges. Generally a harder problem than OOD detection for images
Somewhat casually ignoring the fact that this thesis is mainly interested in audio and speech data, the studies of \cref{part:unsupervised-uncertainty-estimation} dealt with OOD detection in the context of image classification. The reason for this is that image classification has been the de facto standard task for benchmarking OOD detection methods since the first papers on the topic \parencite{choi_waic_2019,nalisnick_detecting_2019,hendrycks_deep_2019,kirichenko_why_2020}. In this section, we shall briefly discuss OOD detection in the context of audio data.

In certain cases, it might be possible to classify an image OOD due to the color or intensity of a single pixel. 
For most distributions, however, we are interested in a test that is sensitive to more semantic properties of the data - for instance the animal class, as discussed in the introduction (\cref{subsec:understanding-uncertainty}). 
For speech, this property holds too; a single amplitude value is unlikely to provide much information about the distributional origin of a speech sample. In fact, audio samples are one-dimensional which likely makes them less discriminatory compared to image pixels that generally have multiple channels e.g. for color. Instead, whether a sample is OOD must be inferred from correlations and patterns found in amplitudes along the temporal dimension.

Few studies have explored OOD detection specifically for speech. One recent study has explored OOD detection for speech \parencite{bukhsh_outofdistribution_2023}. The authors propose a supervised approach that uses a $k$NN classifier fitted to representations of the training data. The approach is tuned to a fixed recall on the training data and operates on full length audio samples. This method is similar to image methods that fit a Gaussian mixture to an appropriate layer \parencite{lee_simple_2018, xiao_we_2021}. In the early work of \textcite{hendrycks_baseline_2017}, the method (based on softmax probabilities) is also benchmarked on audio datasets including TIMIT and Aurora-2 \parencite{garofolo_timit_1993,pearce_aurora_2000}.

% - Harder to work with: Images provide immediate, human-interpretable feedback but audio requires playback which takes time.
The studies on OOD detection for audio reflect its similarity to OOD detection for images. 
Nonetheless, OOD detection on audio comes with unique properties. Where images are immediately interpretable to a researcher, audio requires playback, often in real-time, to present its contents. This fact alone seems likely to play a role in OOD detection being understudied for audio data by simply making it more time-consuming. On top of this, audio is generally variable length and requires more memory than images.

% - Temporal scale, can be OOD in local segments but not entire sequence.

@@

\begin{figure}
    \centering
    \begin{subfigure}[c]{0.49\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{paper_retrospective_calibration_plots/calibration_curve_ensemble_and_all_models_uncalibrated.pdf}
    \end{subfigure}
    \hfill
    \begin{subfigure}[c]{0.49\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{paper_retrospective_calibration_plots/histogram_ensemble_and_single_model.pdf}
    \end{subfigure}
    \caption[Calibration curve for the uncalibrated stroke recognition ensemble and empirical distribution of predicted probabilities.]{ Calibration curve for the uncalibrated stroke recognition ensemble (left) and the histogram of predicted probabilities (right) for the test set. We use the ensemble that achieved the median F1-score reported in \cref{fig_retrospective:figure1-roc-curve} and \cref{tab_retrospective:table3-occlusion-analysis}.}
    \label{fig_discussion:retrospective-paper-calibration-curve-of-uncalibrated-model}
\end{figure}

@@

\begin{figure}
    \centering
    \begin{subfigure}[c]{0.49\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{paper_retrospective_calibration_plots/calibration_curves_ensemble_with_cis.pdf}
    \end{subfigure}
    \hfill
    \begin{subfigure}[c]{0.49\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{paper_retrospective_calibration_plots/calibration_fits_ensemble.pdf}
    \end{subfigure}    
    \caption[Calibration fits and curves for the stroke recognition ensemble using Platt-scaling and isotonic regression for calibration.]{ Calibration curves using sigmoid and isotonic calibration fits for the stroke recognition ensemble model (left) and the calibration fits (right) for the test set. We use the ensemble that achieved the median F1-score reported in \cref{fig_retrospective:figure1-roc-curve} and \cref{tab_retrospective:table3-occlusion-analysis}.}
    \label{fig_discussion:retrospective-paper-calibration-curve-sigmoid-isotonic}
\end{figure}

@@

\begin{figure}
    \centering
    \begin{subfigure}[c]{0.49\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{paper_retrospective_calibration_plots/f1_score_vs_predicted_probability_ensemble_calltaker_with_cis.pdf}
    \end{subfigure}
    \hfill
    \begin{subfigure}[c]{0.49\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{paper_retrospective_calibration_plots/predicted_probability_histogram.pdf}
    \end{subfigure}    
    \caption[Comparison of F1-score of stroke recognition ensemble and call-takers as function of predicted probability.]{ Comparison of the F1-score of the stroke recognition ensemble and call-takers. The F1-score is computed on subsets of the test dataset made by binning on the predicted probabilities of the calibrated ensemble. We see that the relative performance improvement of the ensemble over call-takers is higher towards more certain predictions. We use the ensemble that achieved the median F1-score reported in \cref{fig_retrospective:figure1-roc-curve} and \cref{tab_retrospective:table3-occlusion-analysis}.}
    \label{fig_discussion:retrospective-paper-f1-performance-vs-predicted-probability}
\end{figure}

>>>>

\chapter[discussion]{Discussion}\label{chp:discussion}
% ~5 pages

@@

\moretodo{Adjust the below to include any discussion on out-of-distribution detection.}
Having examined VAEs and self"=supervised approaches for representation learning in \cref{part:unsupervised-uncertainty-estimation,part:unsupervised-speech-representation-learning}, in this discussion we will consider how VAEs might be enabled to learn representations more competitive with self"=supervised methods for downstream tasks (\cref{sec_discussion:representation-learning-with-vaes}). We will also complement the works on OOD detection in \cref{part:unsupervised-uncertainty-estimation}, which focused on image data, with a discussion about similarities with OOD detection for audio and speech data (\cref{sec_discussion:out-of-distribution-detection-for-speech}). 
Furthermore, since the work presented in \cref{part:medical-applications} only had limited focus on uncertainty, we will also present and discuss the use of calibration techniques for the stroke recognition model, drawing connections to how such systems might be perceived and used in practice (\cref{sec: discussion-stroke-recognition-uncertainty}).

@@

\section{Out-of-distribution detection for speech}\label{sec_discussion:out-of-distribution-detection-for-speech}
%
% \moretodo[inline]{Out-of-distribution detection on speech?
%
% Outline:
% Intro paragraph drawing from OOD detection on images.
% Differences and unique challenges. Generally a harder problem than OOD detection for images
Somewhat casually ignoring the fact that this thesis is mainly interested in audio and speech data, the studies of \cref{part:unsupervised-uncertainty-estimation} dealt with OOD detection only in the context of image classification. The reason for this is that image classification has been the de facto standard task for benchmarking OOD detection methods since the first papers on the topic \parencite{choi_waic_2019,nalisnick_detecting_2019,hendrycks_deep_2019,kirichenko_why_2020}. 
In this section, we shall provide some perspective on OOD detection in the context of audio data.

In certain cases, it might be possible to classify an image OOD due to the color or intensity of a single pixel. 
For most distributions, however, we are interested in a test that is sensitive to more semantic properties of the data - for instance the animal class, as discussed in the introduction (\cref{subsec:understanding-uncertainty}). 
For speech, this property holds too; a single amplitude value is unlikely to provide much information about the distributional origin of a speech sample. In fact, audio samples are one-dimensional which likely makes them less discriminatory compared to image pixels that generally have multiple channels e.g. for color. Instead, whether a sample is OOD must be inferred from correlations and patterns found in amplitudes along the temporal dimension. In this way, the task of OOD detection has a similar formulation across different data types.

Few studies have explored OOD detection specifically for speech. One recent study has explored OOD detection for speech \parencite{bukhsh_outofdistribution_2023}. The authors propose a supervised approach that uses a $k$NN classifier fitted to representations of the training data. The approach is tuned to a fixed recall on the training data and operates on full length audio samples. This method is similar to image methods that fit a Gaussian mixture to an appropriate layer \parencite{lee_simple_2018, xiao_we_2021}. In the early work of \textcite{hendrycks_baseline_2017}, the OOD detection method (based on softmax probabilities) is also benchmarked on audio datasets including TIMIT and Aurora-2 \parencite{garofolo_timit_1993,pearce_aurora_2000}. These works strongly suggest that with good data representations, we can use similar methods for OOD detection on both audio and images.

% - Harder to work with: Images provide immediate, human-interpretable feedback but audio requires playback which takes time.
% The studies on OOD detection for audio reflect its similarity to OOD detection for images. 
Despite these similarities of the OOD detection task and its methods across data types, audio comes with unique properties. 
Where images are immediately interpretable to a researcher by visual inspection, audio requires playback, often in real-time, to present its contents. This fact alone is likely to have reduced interest in audio data by simply making it more time-consuming to work with than images. On top of this, audio is generally variable length and requires more memory than images which increases complexity and the amount of compute required to obtain equivalent results compared to images. 
As such, audio-based OOD detection is likely to benefit from standardized frameworks like HuggingFace \parencite{wolf_huggingface_2020,lhoest_datasets_2021} which have made eased working with irregular data types.

% - Temporal scale, can be OOD in local segments but not entire sequence. 
Another fundamental difference between audio and images is that audio is additive; two signals can be added together to form a new, mixed signal that is meaningful in the sense that it could also have been recorded. 
Addition of two images on the other hand produces an image that is not generally representative of what might be photographed. 
This property of images has led to the practice of evaluating OOD detection performance by comparing examples between different datasets (e.g. MNIST versus FashionMNIST). 
The additive property of audio allows complementing this dataset-versus-dataset evaluation with an approach where in-distribution examples are augmented to be OOD by mixing with unknown sources. This was explored by \textcite{hendrycks_baseline_2017} but has not seen a wide exploration.

Finally, most work on OOD detection has focused on an example-centric approach where an entire example is either in- or out-of-distribution. 
For audio, however, it is easy to create examples where local segments are OOD while other parts are not. 
Using the additive property, we could add a loud Morse signal on top of a dialogue to change a segment of in-distribution audio to OOD. 
Similarly, silence might be used as a kind of padding to insert entirely OOD segments in between in-distribution parts without addition and without significant boundary effects. 
For images, the equivalent approach is similar to regularization techniques such as CutMix \parencite{yun_cutmix_2019} which paste images on top of each other. But this usually creates strong boundary errors, which might be used or learned to easily detect such synthetic OOD examples. 
Examples that contain locally OOD parts might occur naturally too; both in the image and audio domains. All kinds of unknown sound can be considered OOD and hence any recording might be locally OOD. In image segmentation, unknown object classes might similarly be seen as locally OOD parts of an example. Future research might examine further how detection of locally OOD examples differs from the dominant example-centric approach.

@@

\begin{figure}[t!]
    \centering
    \begin{subfigure}[c]{0.49\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{paper_retrospective_calibration_plots/calibration_curve_ensemble_and_all_models_uncalibrated.pdf}
    \end{subfigure}
    \hfill
    \begin{subfigure}[c]{0.49\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{paper_retrospective_calibration_plots/histogram_ensemble_and_single_model.pdf}
    \end{subfigure}
    \caption[Calibration curve for the uncalibrated stroke recognition ensemble and empirical distribution of predicted probabilities.]{ Calibration curve for the uncalibrated stroke recognition ensemble (left) and the histogram of predicted probabilities (right) for the test set. We use the ensemble that achieved the median F1-score reported in \cref{fig_retrospective:figure1-roc-curve} and \cref{tab_retrospective:table3-occlusion-analysis}.}
    \label{fig_discussion:retrospective-paper-calibration-curve-of-uncalibrated-model}
\end{figure}

@@

\begin{figure}[t!]
    \centering
    \begin{subfigure}[c]{0.49\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{paper_retrospective_calibration_plots/calibration_curves_ensemble_with_cis.pdf}
    \end{subfigure}
    \hfill
    \begin{subfigure}[c]{0.49\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{paper_retrospective_calibration_plots/calibration_fits_ensemble.pdf}
    \end{subfigure}    
    \caption[Calibration fits and curves for the stroke recognition ensemble using Platt-scaling and isotonic regression for calibration.]{ Calibration curves using sigmoid and isotonic calibration fits for the stroke recognition ensemble model (left) and the calibration fits (right) for the test set. We use the ensemble that achieved the median F1-score reported in \cref{fig_retrospective:figure1-roc-curve} and \cref{tab_retrospective:table3-occlusion-analysis}.}
    \label{fig_discussion:retrospective-paper-calibration-curve-sigmoid-isotonic}
\end{figure}

@@

\begin{figure}[t!]
    \centering
    \begin{subfigure}[c]{0.49\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{paper_retrospective_calibration_plots/f1_score_vs_predicted_probability_ensemble_calltaker_with_cis.pdf}
    \end{subfigure}
    \hfill
    \begin{subfigure}[c]{0.49\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{paper_retrospective_calibration_plots/predicted_probability_histogram.pdf}
    \end{subfigure}    
    \caption[Comparison of F1-score of stroke recognition ensemble and call-takers as function of predicted probability.]{ Comparison of the F1-score of the stroke recognition ensemble and call-takers. The F1-score is computed on subsets of the test dataset made by binning on the predicted probabilities of the calibrated ensemble. We see that the relative performance improvement of the ensemble over call-takers is higher towards more certain predictions. We use the ensemble that achieved the median F1-score reported in \cref{fig_retrospective:figure1-roc-curve} and \cref{tab_retrospective:table3-occlusion-analysis}.}
    \label{fig_discussion:retrospective-paper-f1-performance-vs-predicted-probability}
\end{figure}
